Tree size and hydraulic traits shape drought responses in a temperate broadleaf forest

### 2019-08-14 manuscript/climate_sensitivity_draft.tex
update manuscript, added SI template
changed title

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Tree size, exposure, and hydraulic traits interactively shape drought response in a temperate broadleaf forest},
            pdfauthor={Ian McGregor, Ryan Helcoski, Norbert Kunert, Alan Tepley, Valentine Herrmann, Joseph Zailaa, Atticus Stovall, Neil Pederson, Lawren Sack, Krista Anderson-Teixeira},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{apalike}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

@@

  \title{Tree size, exposure, and hydraulic traits interactively shape drought
response in a temperate broadleaf forest}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Ian McGregor, Ryan Helcoski, Norbert Kunert, Alan Tepley, Valentine
Herrmann, Joseph Zailaa, Atticus Stovall, Neil Pederson, Lawren Sack,
Krista Anderson-Teixeira}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

@@

\usepackage{float}
\usepackage{booktabs}
\usepackage{pdflscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\usepackage{caption}
\captionsetup[table]{font=small}
\captionsetup[figure]{font=small}
\usepackage{dcolumn}

@@

\hypertarget{abstract}{%
\subsubsection{Abstract}\label{abstract}}

\emph{content from AGU abstract:} Here, we integrate forest census data
from a 25.6-ha ForestGEO plot in Virginia (USA), tree-ring records for
12 species representing \#\#\% of woody productivity, leaf hydraulic
trait measurements, and microhabitat data. Individual-level growth
responses to three droughts (1964-66, 1977, 1999) were stronger among
taller trees in dominant canopy positions, those in wetter microsites,
and for more drought-sensitive species as assessed by leaf traits
(turgor loss at less negative leaf water potential, greater shrinkage
with leaf dehydration), with substantial variation in the best predictor
variables across given droughts. We conclude that when droughts occur,
large dominant trees, drought sensitive species, and individuals in
moist microhabitats are likely to be most strongly affected.

\hypertarget{introduction}{%
\subsubsection{Introduction}\label{introduction}}

Understanding how and why trees respond to drought is critical to
predicting forest drought responses and climate change feedbacks.

Forests are diverse in terms of tree sizes and functional traits, and it
is known that trees varying in size and functional traits respond
differently to drought (e.g., \citep{bennett_larger_2015}; REFS).
Therefore, in order to understand whole-forest response to drought, we
need to know how responses vary by tree size/ species. To do so, there
are four fundamental questions that must be addressed:

\emph{First, how significant is the effect of individual drought years?}
Droughts are rarely explicitly defined in ecological studies
\citep{slette_how_2019}, yet no two droughts are the same. This study
addressed this by analyzing trees' resistance to drought within and
across three defined drought periods.

\emph{Second, what drives the observed tendency for large trees to
suffer more during drought?}\\
\citep{bennett_larger_2015} showed that in forests globally, large trees
suffer greater growth reductions during drought. However, this analysis
quantified tree size based on DBH, which has no direct mechanistic
meaning. This study proposed three major mechanisms (besides insects):
(1) inherently greater biophysical challenge of being tall; (2) greater
exposure of the crowns of large trees; and (3) greater water
availability. It is also expected that roots play a role, though these
hypotheses still need to be tested.

\emph{Third, how do species' traits influence drought response?}
Analyzing drought responses on the species level does not fully explain
mechanisms and is not feasible in diverse forests. The solution is a
trait-based approach. Leaf hydraulic traits hold more promise than more
commonly/traditionally-measured traits such as wood density and SLA
(Medeiros et al.).

\emph{Fourth, how do tree size and functional traits interact to
influence drought response?} It is possible that the pattern observed by
\citep{bennett_larger_2015} could be caused by smaller trees being more
drought resistant. Alternatively, larger trees may have more
drought-resistant traits as adpatations to greater biophysical
challenges.

\emph{Hypotheses}

\textbf{1. How significant are individual drought years on a long
time-frame trend?} 1.1 Individual drought scenario is a strong predictor
of drought stress.

\begin{verbatim}
* P1.1 - Drought stress is proportional to the severity of each drought.
\end{verbatim}

\textbf{2. Why do larger trees suffer greater growth declines during
drought?} Our forest displays the same trend as most forests globally
\citep{bennett_larger_2015}. (Note that Bennett et al.~2015 identified
only one study on tree growth responses to drought in the Eastern US
temperate deciduous biome. We know little about how tree size shapes
drought response in this biome.)

2.1 DBH is a strong predictor of drought stress.

\begin{verbatim}
* P2.2-Drought response increases with DBH at time of drought.
\end{verbatim}

2.2. Height is a strong predictor of drought stress.

\begin{verbatim}
* P2.2-Drought response increases with height at time of drought.
* P2.2a - Drought response is better predicted by height than DBH.
\end{verbatim}

2.3. Large trees suffer more during drought because of greater exposure
(to radiation, wind, etc.)--either in relation to neighboring trees or
because of position on landscape.

\begin{verbatim}
* P3.1- Trees currently in a dominant canopy position suffered more during drought. 
\end{verbatim}

2.4. Rooting volume/depth relative to water sources are critical in
drought response. Effects of drought on larger trees are mediated by the
fact that large trees have better access to water.

\begin{verbatim}
 * P3.2- drought response increases with topographic wetness index
\end{verbatim}

\textbf{3. Do species functional traits predict drought response?}

\begin{verbatim}
 * P3.1 - diffuse porous species are more sensitive than ring porous (previously observed in eastern dec forests- Elliot et al. 2015, Friedrichs et al. 2009)
 * P3.2 - higher percent leaf area predicts higher drought stress
 * P3.3 - higher leaf mass area correlates positively to drought resistance (more sclerophyllous vegetation [thick leaves] usually means more adaptation) [Abrams 1990] [Guerfel et al 2009]
 * P3.4 - TLP correlates negatively with drought resistance (NEVER tested),
 * P3.5 - higher wood density means more drought resistant
\end{verbatim}

\textbf{4. How are functional traits distributed across size classes,
and how does this affect size-resistance relationship?}

4.1. Larger/ more exposed trees have more drought resistant traits,
meaning size effects are buffered by traits\\
4.1\_alt. Larger trees suffer more because they have more drought
vulnerable traits.

\begin{verbatim}
 * P3.1a- TLP is lower (larger negative) in taller/canopy trees
 * P3.1b- diffuse porous species more common in understory

 * P3.2- Inclusion of TLP / rp in model does not eliminate or significantly reduce effect of tree size.
\end{verbatim}

\textbf{I think for \#4 here the focus should be on hypotheses looking
at the overall trend compared to the specific scenarios?}

\textbf{4. How do droughts vary in their affect on tree resistance? }

\begin{verbatim}
* P4.1 - The combined-years scenario will have more explaining variables than the individual scenarios.
* P4.2 - Inclusion of leaf hydraulic traits does not eliminate or significantly reduce effect of tree size.
* P4.3 - Drought resistance for each scenario will be explained more by hydraulic traits than by the biophysical environment
\end{verbatim}

\hypertarget{methods}{%
\subsubsection{Methods}\label{methods}}

@@

\emph{Data collection and preparation} The SCBI ForestGEO plot was
censused in 2008, 2013, and 2018 following standard ForestGEO
protocols\ldots{} {[}DETAILS{]}. Here, we use \ldots(\emph{very brief
summary of which data we use for what purpose-- but some will be better
integrated with methods descriptions below})

We analyzed tree-ring data from the twelve species contributing most to
woody aboveground net primary productivity (ANPP\_stem) (Table 1), which
together comprised 97\% of whole-ecosystem ANPP\_stem between 2008 and
2013. Cores were collected in 2010-2011 or 2016-2017 from a breast
height of 1.3m using a 5mm increment borer. In 2010-2011, cores were
collected from randomly selected live trees were selected at random from
each species in 2010-2011, with at least 30 of those trees having a
diameter at breast height (DBH) of at least 10cm
\citep{bourg_initial_2013}. In 2016-2017, cores were collected from all
dead trees found in the annual mortality census
\citep{gonzalezakre_patterns_2016}. Cores were sanded, measured, and
cross-dated using standard procedures, as detailed in
\citep{helcoski_growing_2019}. The resulting chronologies have been
published in association with \citep{helcoski_growing_2019}: (GitHub
URL), (ITRDB).

Height measurements (n=\# trees) were taken by several researchers
between 2012 to 2019, and are archived in a public GitHub repository
(\emph{GitHUB URL}). Measurement methods included manual
\citep[NEON]{stovall_assessing_2018}, digital rangefinders
\citep{andersonteixeira_size-related_2015}, and automatic LiDAR
\citep{stovall_terrestrial_2018}. Rangefinders either used the tangent
method (Impulse 200LR, TruPulse 360R) or the sine method (Nikon
ForestryPro) for calculating heights. The associated errors for using
either method were acknowledged \citep{larjavaara_measuring_2013}.
Species-specific height allometries were developed (Table S\# -
\textbf{ADD THIS TABLE TO SI}). For species that didn't have enough
height measurements, heights were calculated from equations derived from
all species in the study.

@@

\[ diamYEAR = dbh2008 - 2*(bark.depth2008) - 2*\Sigma(ring.widthYEAR:ring.width2008) + 2*(bark.depth_YEAR) \]

@@

Hydraulic traits were collected from SCBI and are summarized in Table 1.
In August 2018, we collected leaf samples from three individuals of each
species \ldots{} (\textbf{Nobby's description of methods for the
following}) 1. PLA 2. LMA 4. Wood density 5. TLP

\textbf{Table 1. Species analyzed here, listed in descending order of
ANPP\_stem. n cores and DBH range represented, and species traits}
{[}*This replaces/combines the two remaining tables in this section.
Suggested columns, with those to include only if they fit in
parentheses: species, (stems \textgreater=10 cm per ha in plot),
(ANPP\_stem), n cores, DBH range of cores, (n cores in each crown
position) species means for each trait{]}

\begin{table}[H]
\centering
\begin{tabular}{l|r|r|r|r|r|r}
\hline
sp & n\_cores & dominant & co-dominant & intermediate & suppressed & prior dead\\
\hline
caco & 13 & NA & 2 & 5 & 5 & 1\\
\hline
cagl & 31 & 1 & 8 & 16 & 5 & 1\\
\hline
caovl & 23 & 4 & 5 & 12 & 2 & NA\\
\hline
cato & 13 & NA & NA & 6 & 2 & 5\\
\hline
fagr & 80 & NA & 7 & 48 & 25 & NA\\
\hline
fram & 62 & NA & 17 & 19 & 14 & 12\\
\hline
juni & 31 & NA & 21 & 8 & NA & 2\\
\hline
litu & 98 & 9 & 29 & 25 & 30 & 5\\
\hline
qual & 61 & 4 & 34 & 20 & 3 & NA\\
\hline
qupr & 59 & 1 & 26 & 20 & 12 & NA\\
\hline
quru & 69 & 6 & 36 & 23 & 2 & 2\\
\hline
quve & 77 & 6 & 46 & 22 & 1 & 2\\
\hline
\end{tabular}
\end{table}
\begin{table}[H]
\centering
\begin{tabular}{l|l|r|r|r}
\hline
Trait & Unit & mean & min & max\\
\hline
Ring Porosity & ring, semi-ring, diffuse & NA & NA & NA\\
\hline
Percent Leaf Area & \% & 15.09 & 8.52 & 24.64\\
\hline
Leaf Mass Area & g/m2 & 53.50 & 30.68 & 75.80\\
\hline
Wood density & g/cm3 & 0.70 & 0.40 & 1.09\\
\hline
TLP & MPa & -2.36 & -2.76 & -1.92\\
\hline
\end{tabular}
\end{table}

@@

\begin{tabular}{rllllllrllrllrllrll}

@@

prediction & variable & variable\_description & null\_model & tested\_model & null\_model\_year & tested\_model\_year & dAIC\_all & coef\_all & coef\_var\_all & dAIC\_1964.1966 & coef\_1964.1966 & coef\_var\_1964.1966 & dAIC\_1977 & coef\_1977 & coef\_var\_1977 & dAIC\_1999 & coef\_1999 & coef\_var\_1999\\

@@

\end{landscape}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(stargazer)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Please cite as:
\end{verbatim}

\begin{verbatim}
##  Hlavac, Marek (2018). stargazer: Well-Formatted Regression and Summary Statistics Tables.
\end{verbatim}

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Tree size and hydraulic traits shape drought responses in a temperate broadleaf forest},
            pdfauthor={Ian McGregor, Ryan Helcoski, Norbert Kunert, Alan Tepley, Valentine Herrmann, Joseph Zailaa, Atticus Stovall, Neil Pederson, Lawren Sack, Krista Anderson-Teixeira},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{apalike}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

@@

  \title{Tree size and hydraulic traits shape drought responses in a temperate
broadleaf forest}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Ian McGregor, Ryan Helcoski, Norbert Kunert, Alan Tepley, Valentine
Herrmann, Joseph Zailaa, Atticus Stovall, Neil Pederson, Lawren Sack,
Krista Anderson-Teixeira}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

@@

\usepackage{float}
\usepackage{booktabs}
\usepackage{pdflscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\usepackage{caption}
\captionsetup[table]{font=small}
\captionsetup[figure]{font=small}
\captionsetup[table]{labelformat=empty}
\usepackage{dcolumn}

@@

\subsubsection{Summary}\label{summary}

Predicting forest responses to drought is an increasingly critical task
under climate change effects. Part of the problem is due to the lack of
studies analyzing the confluence of leaf hydraulic traits with
biophysical parameters. In this study, we analyze the interaction
between these two trait groups using forest census data from a 25.6-ha
ForestGEO plot in Virginia (USA). Drought periods were defined by both
Palmer Drought Severity Indices (PDSI) and their identification from
tree-ring records for 12 species representing 97\% of woody
productivity. Each drought scenario (1964-66, 1977, 1999), along with
the overall trend, was then tested against leaf hyraulic trait
measurements and microhabitat biophysical data. Individual-level growth
responses to the three individual droughts were stronger among taller
trees in dominant canopy positions, those in wetter microsites, and for
more drought-sensitive species as assessed by leaf traits (turgor loss
at less negative leaf water potential, greater shrinkage with leaf
dehydration), with substantial variation in the best predictor variables
across given droughts. We conclude that when droughts occur, large
dominant trees, drought-sensitive species, and individuals in wetter
microhabitats are likely to be most strongly affected.

\emph{The Summary for research papers, which must be usable as a stand-
alone document, must not exceed 200 words and should be organized using
four bullet points to indicate: (1) the research conducted, including
the rationale, (2) methods, (3) key results, and (4) the main
conclusion, including the key points of discussion. It should not
contain citations of other papers.}

\subsubsection{Introduction}\label{introduction}

Understanding how and why trees respond to drought is critical to
predicting forest drought responses and climate change feedbacks. (1
paragraph on this- I have lots of content on this and will add later.)

{[}Understanding forest responses to drought requires increased
functional understanding of the factors that confer individual-level
vulnerability or resistance.{]} Forests are diverse in terms of tree
sizes and functional traits, and it is known that trees varying in size
and functional traits respond differently to drought (e.g.,
\citep{bennett_larger_2015}; REFS). Therefore, in order to understand
whole-forest response to drought, we need a functional understanding of
how responses vary by tree size, microhabitat, and species. There are 3
fundamental questions that must be addressed:

\emph{First, what drives the observed tendency for large trees to suffer
more during drought?}\\
\citet{bennett_larger_2015} showed that in forests globally, large trees
suffer greater growth reductions during drought. However, this analysis
quantified tree size based on DBH, which has no direct mechanistic
meaning. This study proposed two major mechanisms (besides insects) for
the observed greater drougth growth reductions of large trees: (1)
inherently greater biophysical challenge of being tall; (2) greater
exposure of the crowns of large trees. Counteracting these effects, (3)
the larger root systems of larger trees may confer an advantage in terms
of allowing greater access to water, but it appears that this effect is
usually insufficient to offset the costs of height and/or crown
exposure.

Canopy trees have lower drought resistance because they are exposed to
higher solar radiation, greater wind speeds, and lower humidity.
Alternatively, the generally supressed status of subcanopy trees may be
insufficient to override the benefits of their buffered environment
during drought.

\emph{Second, how do species' traits - alone and in interaction with
tree size - influence drought response?} Analyzing drought responses on
the species level does not fully explain mechanisms and is not feasible
in diverse forests. The solution is a trait-based approach. Leaf
hydraulic traits hold more promise than more
commonly/traditionally-measured traits such as wood density and specific
leaf area (SLA)
(\href{https://besjournals.onlinelibrary.wiley.com/doi/abs/10.1111/1365-2435.13229}{Medeiros
et al. 2019}).

Commonly measured traits including wood density (REFS), leaf mass per
area \citep{abrams_adaptations_1990} \citep{guerfel_impacts_2009}, and
ring porosity (Elliot et al. 2015, Friedrichs et al. 2009) have been
linked to drought responses and likely correlated with drought
resistance in this forest. However, we hypothesize that leaf hydraulic
traits such as leaf area shrinkage upon dessication (PLA) and turgor
loss point (TLP), which are emerging as potentially more informative
traits but whose effect on drought resistance has never been tested
(CONFIRM), will prove better predictors.

Moreover, there may be an interaction between traits and tree size. It
is possible that the pattern observed by \citet{bennett_larger_2015}
could be caused by smaller trees being more drought resistant.
Alternatively, larger trees may have more drought-resistant traits as
adpatations to greater biophysical challenges.

\emph{Third, are responses similar or variable across individual drought
years?} Droughts are rarely explicitly defined in ecological studies
\citep{slette_how_2019}, yet no two droughts are the same.

We need to understand the the factors confirming drought vulnerability
or resistance not only for extreme droughts with dramatic impacts on
tree growth and mortality, which tend to dominate the literature
\citep{bennett_larger_2015} (REFS), but also for more modest but
frequent droughts--e.g., those with historical return intervals on the
order of a decade.

Here, we combine tree-ring records covering three droughts (1964-66,
1977, 1999), species functional and hydraulic trait measurments, and
forest census data from a 25.6-ha ForestGEO plot in Virginia (USA) to
test a series of hypotheses and associated specific predictions (Table
1) designed to yield functional understanding of how tree size,
microenrivonment, and species' traits collectively shape drought
responses. First, we focus on the role of tree size and its interaction
with microenvironement. We confirm that, consistent with most forests
globally \citep{bennett_larger_2015}, larger-diameter trees have lower
drought resistance in this forest, which is in an ecoregion represented
by only one study in Bennett et al (2015) (H1.0). We then test
hypotheses designed to disentangle the relative importance of tree
height (H1.1), crown exposure (H1.2), and root water access, which
should be greater for larger trees in dry but not perpetually wet
microsites (H1.3). Second, we focus on the role of species' functional
and hydraulic traits and their interaction with tree height. We
hypothesize that drought resistance will follow predicted and observed
patterns in relation to wood density, specific leaf area, and ring
porosity (P2.1a-c), but that leaf hydraulic traits such as leaf area
shrinkage upon dehydration and turgor loss point will prove better
predictors (P2.1d-e). We then test whether these traits correlate with
tree height (P2.2), potentially driving the observed tendency for taller
trees to suffer more during drought (P2.3). Finally, we focused on
variability among droughts, asking how community resistance varied
across droughts (H3.1) and whether the factors confirming vulnerability
or resistance varied across droughts (H3.2).

\emph{Let's put Table 1 here. We need to give the hypotheses/
predictions here. It seems counterintuitive to put the results right up
front like that, but it actually tends to be a higher-impact style of
writing, and recommended
\href{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1005619}{here}.}

\textbf{Table 1. Summary of hypotheses, predictions and results}
\includegraphics[width=5.20833in]{tables_figures/Table1.png}

We count predictions as fully supported (or rejected) when the direction
of response matches (or contradicts) the prediction. H2.3-H3.2 are based
on if models containing species traits in H2.1 had dAIC\textgreater{}=2
relative to the appropriate null model or to any alternative
multivariate model within 2 dAIC. Parentheses indicate that predictions
were partially supported (or rejected). In other words, parentheses
indicate when the direction of response matched (or contradicted) the
prediction in some but not all models with dAIC\textless{}2 relative to
the appropriate null model. With categorical variables such as crown
position, a ``(yes)'' notation describes when the trend matched the
prediction, but it wasn't significant A ``(yes/no) or (''no/yes``)''
indicates tendencies in opposite directions in univariate tests versus
the best full models, respectively. ``(yes)

\subsubsection{Materials and Methods}\label{materials-and-methods}

@@

\emph{Data collection and preparation} The SCBI ForestGEO plot was
censused in 2008, 2013, and 2018 following standard ForestGEO protocols,
whereby all free-standing woody stems \textgreater{}= 1cm diameter at
breast height (DBH) were mapped, tagged, measured at DBH, and identified
to species \citep{condit_tropical_1998}. From this census data, we used
measurements of DBH from 2008 to calculate historical DBH (described
below), along with tree location in the plot to determine the
topographic wetness index. Furthermore, we used data for all stems
\textgreater{}= 10cm to analyze functional trait composition relative to
tree height. \emph{I added the last sentence-- not sure if you've
described that below yet. I don't think you need to note here which data
we use, so long as that's described in the portions where you describe
the analysis.}

We analyzed tree-ring data from the twelve species contributing most to
woody aboveground net primary productivity (ANPP), which together
comprised 97\% of study plot ANPP between 2008 and 2013
\citep{helcoski_growing_2019}. Cores were obtained in 2010-2011 or
2016-2017 from a breast height of 1.3m using a 5mm increment borer. In
2010-2011, cores were collected from randomly selected live trees of
species with at least 30 individuals of DBH \textgreater{}= 10cm
\citep{bourg_initial_2013}. In 2016-2017, cores were collected from all
trees found dead in the annual mortality census
\citep{gonzalezakre_patterns_2016}. Cores were sanded, measured, and
cross-dated using standard procedures, as detailed in
\citep{helcoski_growing_2019}. The resulting chronologies have been
published in association with \citet{helcoski_growing_2019}: (ITRDB;
GitHub/Zenodo). \emph{is this a database? Yes. Ryan submitted the data
but I don't think its posted yet. We should also cite GitHub/Zenodo
here. I'll come back to that. }

Height measurements (n=\# trees) were taken by several researchers
between 2012 to 2019, and are archived in a public
\href{https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_dimensions/tree_heights}{GitHub
repository}. Measurement methods included manual
\citep[NEON]{stovall_assessing_2018}, digital rangefinders
\citep{andersonteixeira_size-related_2015}, and automatic LiDAR
\citep{stovall_terrestrial_2018}. Rangefinders either used the tangent
method (Impulse 200LR, TruPulse 360R) or the sine method (Nikon
ForestryPro) for calculating heights. The associated errors for using
either method were acknowledged \citep{larjavaara_measuring_2013}.
Species-specific height allometries were developed (Table S\# -
\textbf{ADD THIS TABLE TO SI}). For species that didn't have enough
height measurements, heights were calculated from equations derived from
all species in the study.

@@

\[ diamYEAR = dbh2008 - 2*(bark.depth2008) - 2*\Sigma(ring.widthYEAR:ring.width2008) + 2*(bark.depthYEAR) \]

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=5.20833in]{tables_figures/Figure3.png}
\caption{Location of cored trees}
\end{figure}

Hydraulic traits were collected from SCBI and are summarized in Table 1.
In August 2018, we collected leaf samples from three individuals of each
species \ldots{} (\textbf{Nobby's description of methods for the
following (see word document}) 1. PLA 2. LMA 4. Wood density 5. TLP

\textbf{Table 2. Species analyzed here, listed in descending order of
ANPP\_stem. n cores and DBH range represented, and species traits}
{[}*This replaces/combines the two remaining tables in this section.
Suggested columns, with those to include only if they fit in
parentheses: species, (stems \textgreater{}=10 cm per ha in plot),
(ANPP\_stem), n cores, DBH range of cores, (n cores in each crown
position) species means for each trait{]}

\begin{verbatim}
## Warning: package 'kableExtra' was built under R version 3.4.4
\end{verbatim}

\textbackslash{}begin\{table\}{[}!h{]}

\textbackslash{}caption\{\label{tab:Table2}Overview of analyzed species,
detailing DBH mean and range of cored trees, the number of cores
represented by each crown position of each species, and mean hydraulic
trait measurements. Units of measurements are in mm (DBH), \% (PLA),
g/m2 (LMA), MPa (TLP), and g/cm3 (WD).\} \centering

\begin{tabular}{lrrlrrrr}
\toprule
sp & mean\_DBH & range\_DBH & RP & PLA & LMA & TLP & WD\\
\midrule
caco & 271.87 & 508.0 & ring & 17.22 & 45.86 & -2.13 & 0.83\\
cagl & 313.89 & 887.0 & ring & 21.09 & 42.76 & -2.13 & 0.62\\
caovl & 352.87 & 511.0 & ring & 14.80 & 47.60 & -2.48 & 0.96\\
cato & 209.74 & 201.1 & ring & 16.56 & 45.36 & -2.20 & 0.83\\
fagr & 235.11 & 960.0 & diffuse & 9.45 & 30.68 & -2.57 & 0.62\\
\addlinespace
fram & 353.63 & 883.3 & ring & 13.06 & 43.28 & -2.10 & 0.56\\
juni & 481.42 & 628.0 & semi-ring & 24.64 & 72.13 & -2.76 & 1.09\\
litu & 368.54 & 904.0 & diffuse & 19.56 & 46.92 & -1.92 & 0.40\\
qual & 471.51 & 677.0 & ring & 8.52 & 75.80 & -2.58 & 0.61\\
qupr & 422.48 & 767.0 & ring & 11.75 & 71.77 & -2.36 & 0.61\\
\addlinespace
quru & 548.79 & 1369.3 & ring & 11.01 & 71.13 & -2.64 & 0.62\\
quve & 541.38 & 981.8 & ring & 13.42 & 48.69 & -2.39 & 0.65\\
\bottomrule
\end{tabular}

\textbackslash{}end\{table\}

@@

\begin{tabular}{lrll}

@@

Modnames & Delta\_AICc & scenario & coef\\

@@

\includegraphics{tables_figures/FigureS1.png} (see Issue \#32)
